\documentclass[11pt]{article}
\usepackage{homework}

\classname{322}
\homeworknum{3}



\begin{document}

\input{definitions}
\input{probdefs}

\input{Problem1}

\input{Problem2}



\state{(Jackson 12.19)}{
	Source-free electromagnetic fields exist in a localized region of space.  Consider the various conservation laws that are contained in the integral of $\ptsa \Mabg = 0$ over all space, where %$\Mabg$ is defined by (12.117).
	\eq{
		\Mabg = \Thab \xg - \Thag \xb.
	}
	\vfix
}

%
%	12.19(a)
%

\prob{}{
	Show that when $\bet$ and $\gam$ are both space indices conservation of the total field angular momentum follows.
}

\sol{
	Note that
	\eq{
		\ptsa \Mabg = \ptsa(\Thab \xg - \Thag \xb)
		= \xg \ptsa \Thab + \Thab \ptsa \xg - \xb \ptsa \Thag - \Thag \ptsa \xb
		= \Thab \ptsa \xg - \Thag \ptsa \xb,
	}
	where we have used Jackson~(12.116), $\ptsa \Thab = 0$.
	\clearpage
}

%
%	12.19(b)
%

\clearpage
\prob{}{
	Show that when $\bet = 0$ the conservation law is
	\eq{
		\dv{\vX}{t} = \frac{c^2 \vPem}{\Eem},
	}
	where $\vX$ is the coordinate of the center of mass of the electromagnetic fields, defined by
	\eq{
		\vX \int u \dcx = \int \vx u \dcx,
	}
	where $u$ is the electromagnetic energy density and $\Eem$ and $\vPem$ are the total energy and momentum of the fields.
}


\clearpage
\state{}{
	We discussed in class the construction of linearly polarized electromagnetic waves.
}

%
%	4(a)
%

\prob{}{
	Generalize the discussion to circularly polarized waves (see also Wald Sec.~{5.5}).  Discuss both right-handed and left-handed polarizations.
}

\sol{
	The plane waves are given on p.~149--150 in the lecture notes, and by Eq.~(7.8) in Jackson:
	\al{
		\vE(\vr, t) &= \vsE \exp(i \ksm \xm), &
		\vB(\vr, t) &= \vsB \exp(i \ksm \xm),
	}
	where $\kkm = (\omg / c, \vk)$, with $\omg$ being the wave frequency and $\vk$ the wave vector.  Also, from Jackson~(7.11--12),
	\al{
		\vsE &= \sE \vxiq, &
		\vsB &= \sE \vxiw
		= \kh \cross \vsE,
	}
	where $\sE$ and $\sB$ with $\sE = \sB$ are the field strength amplitudes, and $\vxiq$ and $\vxiw$ are polarization vectors which are related to the polarization 4-vectors $\ximq$ and $\ximw$.  From p.~146 in the lecture notes, the polarization vectors must both satisfy the constraint $\ksm \xim = 0$ and the identification $\xism \sim \xism + \alp \ksm$, where $\alp$ is an arbitrary constant.  We found in class that this means $\ximq = (0, \vxiq)$ and $\ximw = (0, \vxiw)$.  Then we must have~\cite[p.~299]{Jackson}
	\al{
		\vxiq \vdot \kh &= \vxiw \vdot \kh = \vxiq \vdot \vxiw = 0, &
		\vxiw &= \kh \cross \vxiq,
	}
	where $\kh$ is a unit vector in the direction of $\vk$.
	
	Circularly polarized waves are the linear combination of two linearly-polarized waves that have the same amplitude and are out of phase by $\pi / 2$.  The circularly-polarized fields may then be written as~\cite[p.~299]{Jackson}
	\aln{ \label{circ1}
		\ans{ \vE(\vr, t)\ }&\ans{ = \sE (\vxiq \pm i \vxiw) \exp(i \ksm \xm), } &
		\ans{ \vB(\vr, t)\ }&\ans{ = \sB (\vxiw \mp i \vxiq) \exp(i \ksm \xm), }
	}
	where the upper signs correspond to left-handed polarization, and the lower signs correspond to right-handed polarization.  Incoming waves with left-handed polarization appear to be rotating counter-clockwise to an observer, while incoming right-handed waves appear to be rotating clockwise~\cite[p.~300]{Jackson}.
	
	Taking the real part of Eq.~\refeq{circ1}, and choosing $\vxiq = \xh$, $\vxiw = \yh$, and $\kh = \zh$, we have~\cite[p.~299]{Jackson}
	\aln{
		\ans{ \Ex(z, t)\ }&\ans{ = \sE \cos(k z - \omg t), }&
		\ans{ \Ey(z, t)\ }&\ans{ = \mp \sE \sin(k z - \omg t), } \label{circ2.1} \\
		\ans{ \Bx(z, t)\ }&\ans{ = \pm \sB \sin(k z - \omg t), } &
		\ans{ \By(z, t)\ }&\ans{ = \sB \cos(k z - \omg t). } \label{circ2.2}
	}
	We will show that these fields satisfy Maxwell's equations for a source-free region.  From Wald~(5.4--7),
	\al{
		\grad \vdot \vE &= 0, &
		\grad \cross \vB - \frac{1}{c} \pdv{\vE}{t} &= \vo, &
		\grad \vdot \vB &= 0, &
		\grad \cross \vE + \frac{1}{c} \pdv{\vB}{t} &= \vo.
	}
	For the first equation,
	\eq{
		\grad \vdot \vE = \pdv{\Ex}{x} + \pdv{\Ey}{y}
		= 0.
	}
	For the second,
	\al{
		\grad \cross \vB - \frac{1}{c} \pdv{\vE}{t} &= -\pdv{\By}{z} \,\xh + \pdv{\Bx}{z} \,\yh + \paren{ \pdv{\By}{x} - \pdv{\Bx}{y} } \zh - \frac{1}{c} \paren{ \pdv{\Ex}{t} \,\xh + \pdv{\Ey}{t} \,\yh } \\
		&= \sB k \sin(k z - \omg t) \,\xh \pm \sB k \cos(k z - \omg t) \,\yh - \frac{\sE \omg \sin(k x - \omg t) \,\xh \pm \sE \omg \cos(k x - \omg t) \,\yh}{c}
		= \vo,
	}
	since $\sE = \sB$ in Gaussian units and $k = \omg / c$ from p.~138.
	
	For the third,
	\eq{
		\grad \vdot \vB = \pdv{\Bx}{x} + \pdv{\By}{y}
		= 0.
	}
	For the fourth,
	\al{
		\grad \cross \vE + \frac{1}{c} \pdv{\vB}{t} &= -\pdv{\Ey}{z} \,\xh + \pdv{\Ex}{z} \,\yh + \paren{ \pdv{\Ey}{x} - \pdv{\Ex}{y} } \zh + \frac{1}{c} \paren{ \pdv{\Bx}{t} \,\xh + \pdv{\By}{t} \,\yh } \\
		&= \pm \sE k \cos(k z - \omg t) \,\xh - \sE k \sin(k z - \omg t) \,\yh - \frac{\pm \sB \omg \cos(k z - \omg t) \,\xh - \sB \omg \sin(k x - \omg t)\,\yh}{c}
		= \vo.
	}
	\ans{ Thus, we have shown that the circularly-polarized waves given by Eqs.~(\ref{circ1}--\ref{circ2.2}) are valid free-field solutions to the Maxwell equations. }
}

%
%	4(b)
%

\prob{}{
	Compute the angular momentum of the circularly polarized waves of part~(a) using the formula for angular momentum derived in class.
}

\sol{
	The angular momentum of an electric field is given on p.~128 of the lecture notes as
	\eq{
		\vL = \frac{1}{4\pi c} \int \vr \cross (\vE \cross \vB) \dcx,
	}
	where $\vr = (x, y, z)$ denotes a point in space.
}


\clearpage 
\input{Problem5}


\makebib

\end{document}