\documentclass[11pt]{article}
\usepackage{geometry, titlesec}
\usepackage[parfill]{parskip}
\usepackage[italicdiff]{physics}
\usepackage{amsfonts, amsthm}
\usepackage[cm]{fullpage}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{xcolor, soul}
\usepackage{kbordermatrix}
%\allowdisplaybreaks

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vfix}{\vspace{-\baselineskip}}

\makeatletter
\renewcommand*\env@cases[1][1.2]{%
  \let\@ifnextchar\new@ifnextchar
  \left\lbrace
  \def\arraystretch{#1}%
  \array{@{}l@{\quad}l@{}}%
}
\makeatother
 
\renewcommand{\footrulewidth}{.2pt}
%\setlist[enumerate]{leftmargin=*}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textbf{Physics 316 Homework 7}}
\rhead{Lacey Rainbolt}
\setlength{\headheight}{11pt}
\setlength{\headsep}{11pt}
\setlength{\footskip}{24pt}
\lfoot{\today}
\rfoot{\thepage}

\titleformat{\section}[runin]{\normalfont\large\bfseries}{Problem \thesection.}{1em}{}
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subparagraph}[leftmargin]{\normalfont\normalsize\bfseries}{}{0pt}{}

\newcommand{\refeq}[1]{(\ref{#1})}

\newcommand{\beq}{\begin{equation*}}
\newcommand{\eeq}{\end{equation*}}

\newcommand{\beqn}{\begin{equation}}
\newcommand{\eeqn}{\end{equation}}

\newcommand{\blg}{\begin{align*}}
\newcommand{\elg}{\end{align*}}


\newenvironment{statement}[1]
{
	\section{#1}
	\color{darkgray}
	\ignorespaces
}
{
%    \smallskip
}

\newenvironment{problem}
{
	\subsection{}
	\color{darkgray}
%	\paragraph{Problem.}
    \ignorespaces
}
{

}

\newenvironment{solution}
{
    \paragraph{Solution.}
    \ignorespaces
}
{
    \bigskip
}



\begin{document}

\input{Problem1}

\input{Problem2}

\input{Problem3}

\input{Problem4}


%\newcommand{\gam}{\gamma}
\newcommand{\gami}{\gam^{(i)}}
\newcommand{\gamj}{\gam^{(j)}}
\newcommand{\vr}{\vec{r}}
\newcommand{\vri}{\vr_i}
\newcommand{\vrj}{\vr_j}
\renewcommand{\xi}{x_i}
\newcommand{\yi}{y_i}
\newcommand{\xj}{x_j}
\newcommand{\yj}{y_j}

\newcommand{\xd}{\dot{x}}
\newcommand{\yd}{\dot{y}}
\newcommand{\xdi}{\xd_i}
\newcommand{\ydi}{\yd_i}

\newcommand{\gamq}{\gam^{(1)}}
\newcommand{\gamw}{\gam^{(2)}}
\newcommand{\vrq}{\vr_1}
\newcommand{\vrw}{\vr_2}

%\newcommand{\xq}{x_q}
\newcommand{\yq}{y_1}
\newcommand{\xw}{x_2}
\newcommand{\yw}{y_2}

\newcommand{\xdq}{\xd_1}
\newcommand{\ydq}{\yd_1}
\newcommand{\xdw}{\xd_2}
\newcommand{\ydw}{\yd_2}

\begin{statement}{Interacting line vortices} \label{vort1}
\subparagraph{}
	A system of $n$ vortices moving on a two-dimensional plane has the Hamiltonian
	\beq
		H = \sum_{j = 1}^n \sum_{i = 1}^{j - 1} -\gami \gamj \ln|\vri - \vrj|,
	\eeq
	where $\gami$ is the strength of the $i$th line vortex, and $\vri = (\xi, \yi)$ its position in the plane.  Using the Poisson bracket structure
	\beq
		[f, g] = \sum_{i = 1}^n \frac{1}{\gami} \left( \pdv{f}{\xi} \pdv{g}{\yi} - \pdv{f}{\yi} \pdv{g}{\yi} \right),
	\eeq
	Hamilton's equations for the vortex system simplify to
	\begin{align*}
		\xdi &= [\xi, H], &
		\ydi = [\yi, H].
	\end{align*}
	Consider two vortices.  Show that the equations of motion can be solved explicitly.  Most importantly, show that the solution tells us the two vortices orbit each other with a frequency that is inversely proportional to the square of their separation.
\end{statement}

\begin{solution}
	For two vortices, the Hamiltonian reduces to
	\beq
		H = -\gamq \gamw \ln|\vrq - \vrw|.
	\eeq
	For $i, j \in \{ 1, 2 \}$, note that
	\begin{align*}
		\pdv{\xi}{\xj} &= \pdv{\yi}{\yj} = \delta_{ij}, &
		\pdv{\xi}{\yj} &= \pdv{\yi}{\xj} = 0.
	\end{align*}
	Note also that
	\beq
		|\vrq - \vrw| = \sqrt{(\xq - \xw)^2 + (\yq - \yw)^2}
		= \sqrt{\xq^2 - 2 \xq \xw + \xw^2 + \yq^2 - 2 \yq \yw + \yw^2}.
	\eeq
	Define
	\begin{align*}
		w &\equiv (\xq - \xw)^2 + (\yq - \yw)^2, &
		v &\equiv \sqrt{w}, &
		u &\equiv \ln{v}.
	\end{align*}
	Then
	\begin{align*}
		\pdv{H}{\xi} &= \pdv{H}{u} \pdv{u}{v} \pdv{v}{w} \pdv{w}{\xi}
		= -\gami \gamj \frac{1}{v} \frac{1}{2v} (2 \xi - 2 \xj)
		= -\gami \gamj \frac{\xi - \xj}{(\xq - \xw)^2 + (\yq - \yw)^2}
		= -\gami \gamj \frac{\xi - \xj}{|\vrq - \vrw|^2}, \\
		\pdv{H}{\yi} &= \pdv{H}{u} \pdv{u}{v} \pdv{v}{w} \pdv{w}{\yi}
		= -\gami \gamj \frac{\yi - \yj}{|\vrq - \vrw|^2},
	\end{align*}
	where $i \neq j$.
	
	Combining the above, and again fixing $i \neq j$,
	\begin{align*}
		\xdi &= \frac{1}{\gami} \left( \pdv{\xi}{\xi} \pdv{H}{\yi} - \pdv{\xi}{\yi} \pdv{H}{\xi} \right) + \frac{1}{\gamj} \left( \pdv{\xi}{\xj} \pdv{H}{\yj} - \pdv{\xi}{\yj} \pdv{H}{\xj} \right)
		= \frac{1}{\gami} \pdv{H}{\yi}
		= -\gamj \frac{\yi - \yj}{|\vrq - \vrw|^2}, \\
		\ydi &= \frac{1}{\gami} \left( \pdv{\yi}{\xi} \pdv{H}{\yi} - \pdv{\yi}{\yi} \pdv{H}{\xi} \right) + \frac{1}{\gamj} \left( \pdv{\yi}{\xj} \pdv{H}{\yj} - \pdv{\yi}{\yj} \pdv{H}{\xj} \right)
		= -\frac{1}{\gami} \pdv{H}{\xi}
		= \gamj \frac{\xi - \xj}{|\vrq - \vrw|^2}.
	\end{align*}
	Explicitly, the four equations of motion are
	\begin{align*}
		\xdq &= -\gamw \frac{\yq - \yw}{|\vrq - \vrw|^2}, &
		\xdw &= \gamq \frac{\yq - \yw}{|\vrq - \vrw|^2}, &
		\ydq &= \gamw \frac{\xq - \xw}{|\vrq - \vrw|^2}, &
		\ydw &= -\gamq \frac{\xq - \xw}{|\vrq - \vrw|^2}.
	\end{align*}
\end{solution}


\newcommand{\Px}{P_x}
\newcommand{\Py}{P_y}

\begin{statement}{Conserved quantities for a system of line vortices} \label{vort2}
\subparagraph{}
	Now consider the general case of $n$ vortices.
\end{statement}

\begin{problem}
	Verify that the ``total linear momentum along $x$'' and the ``total linear momentum along $y$,''
	\begin{align*}
		\Px &= \sum_{i = 1}^n \gami \yi, &
		\Py &= \sum_{i = 1}^n -\gami \xi,
	\end{align*}
	are conserved.
\end{problem}

\begin{problem}
	Verify that $[\Px, \Py]$ gives a conserved quantity.
\end{problem}


\newcommand{\vB}{\vec{B}}
\newcommand{\vA}{\vec{A}}
\newcommand{\px}{p_x}
\newcommand{\py}{p_y}
\newcommand{\pdx}{\dot{p}_x}
\newcommand{\pdy}{\dot{p}_y}

\begin{statement}{Charged particle in a magnetic field}
\subparagraph{}
	Suppose a charged particle moves in a two-dimensional plane while experiencing a magnetic field $\vB = (0, 0, B)$.  Use the vector potential $\vA = (-By, 0, 0)$.  The Hamiltonian for the particle is
	\beq
		H = \frac{1}{2m} \left( \px + \frac{e B}{c} y \right)^2 + \frac{\py^2}{2m}.
	\eeq
\vfix
\end{statement}

\begin{problem}
	Write down Hamilton's equations.  Verify that by appropriate manipulation we have
	\begin{align*}
		\py + \frac{e B}{c} x &= a, &
		\px &= m \xd - \frac{e B}{c} y = b,
	\end{align*}
	where $a$ and $b$ are constants.
\end{problem}

\begin{solution}
	Note that
	\beq
		H = \frac{1}{2m} \left( \px^2 + 2 \frac{e B}{c} \px y + \frac{e^2 B^2}{c^2} y^2 \right) + \frac{\py^2}{2m}
		= \frac{\px^2}{2m} + \frac{e B}{c} \frac{\px y}{m} + \frac{e^2 B^2}{c^2} \frac{y^2}{2m} + \frac{\py^2}{2m}
	\eeq
	Hamilton's equations are
	\begin{align}
		\xd &= \pdv{H}{\px} = \frac{1}{m} \left( \px + \frac{e B}{c} y \right), \label{xd} \\
		\pdx &= -\pdv{H}{x} = 0, \label{pdx} \\
		\yd &= \pdv{H}{\py} = \frac{\py}{m}, \label{yd} \\
		\pdy &= -\pdv{H}{y} = -\frac{e B}{c} \frac{1}{m} \left( \px + \frac{e B}{c} y \right). \label{pdy}
	\end{align}
	
	Substituting \refeq{xd} into \refeq{pdy},
	\beq
		\pdy = -\frac{e B}{c} \xd.
	\eeq
	By integrating with respect to $t$, we obtain
	\beq
		\py = -\frac{e B}{c} \int \xd \dd{t} = -\frac{e B}{c} x + a,
	\eeq 
	where $a$ is some constant.  Therefore, we have
	\beqn \label{a}
		\py + \frac{e B}{c} x = a,
	\eeqn
	as desired.
	
	From \refeq{xd},
	\beq
		m \xd = \px + \frac{e B}{c} y
		\iff
		\px = m \xd - \frac{e B}{c} y,
	\eeq
	and from \refeq{pdx},
	\beq
		\px = \int 0 \dd{t} = b,
	\eeq
	where $b$ is some constant.  Combining these, we have
	\beqn \label{b}
		\px = m \xd - \frac{e B}{c} y = b
	\eeqn
	as desired. \qed
\end{solution}

\newcommand{\omg}{\omega}
\newcommand{\xdd}{\ddot{x}}
\newcommand{\ydd}{\ddot{y}}

\newcommand{\xc}{x_c}
\newcommand{\yc}{y_c}
\newcommand{\xddc}{\ddot{x}_c}
\newcommand{\yddc}{\ddot{y}_c}

\newcommand{\xxt}{x(t)}
\newcommand{\yyt}{y(t)}
\newcommand{\xct}{\xc(t)}
\newcommand{\yct}{\yc(t)}
\newcommand{\xpt}{x_p(t)}
\newcommand{\ypt}{y_p(t)}

\newcommand{\Cq}{C_1}
\newcommand{\Cw}{C_2}
\newcommand{\Dq}{D_1}
\newcommand{\Dw}{D_2}

\begin{problem}
	Using the relations above and the equations of motion, verify that the charged particle moves in a circle in the $(x, y)$ plane and that the circling frequency $\omg$ is given by
	\beq
		\omg = \frac{e B}{m c}.
	\eeq
This is called the \emph{Larmor frequency}.
\end{problem}

\begin{solution}
	Substituting \refeq{b} into \refeq{xd} yields
	\beqn \label{xd2}
		\xd = \frac{1}{m} \left( b + \frac{e B}{c} y \right).
	\eeqn
	Similarly, solving \refeq{a} for $\py$ and substituting into \refeq{yd} gives us
	\beqn \label{yd2}
		\yd = \frac{1}{m} \left( a - \frac{e B}{c} x \right).
	\eeqn
	\refeq{xd2} and \refeq{yd2} are a system of two coupled first-order equations.  Differentiating each by $t$, we obtain two uncoupled second-order equations:
	\begin{align*}
		\xdd &= \frac{1}{m} \frac{e B}{c} \yd
		= \frac{1}{m^2} \frac{e B}{c} \left( a - \frac{e B}{c} x \right), &
		\ydd &= -\frac{1}{m} \frac{e B}{c} \xd
		= -\frac{1}{m^2} \frac{e B}{c} \left( b + \frac{e B}{c} y \right).
	\end{align*}
	These are inhomogeneous equations, meaning that their general solutions are each the sum of a complementary solution and a particular solution.  That is,
	\begin{align*}
		\xxt &= \xct + \xpt, &
		\yyt &= \yct + \ypt,
	\end{align*}
	where $\xct$ and $\yct$ are the complementary solutions, and $\xpt$ and $\ypt$ are the particular solutions.
	
	In order to show circular motion, it is sufficient to show that the complementary solutions $\xct$ and $\yct$ respresent circular motion.  (The particular solutions have the same form as the complementary solutions, but with coefficients determined by the constants $a$ and $b$.)  The complementary solutions satisfy
	\begin{align*}
		\xddc &= -\frac{1}{m^2} \frac{e^2 B^2}{c^2} \xc, &
		\yddc &= -\frac{1}{m^2} \frac{e^2 B^2}{c^2} \yc,
	\end{align*}
	which have solutions
	\begin{align*}
		\xct &= \Cq \cos(\frac{e B}{m c} t) + \Cw \sin(\frac{e B}{m c} t)
		= \Cq \cos(\omg t) + \Cw \sin(\omg t), \\
		\yct &= \Dq \cos(\frac{e B}{m c} t) + \Dw \sin(\frac{e B}{m c} t)
		= \Dq \cos(\omg t) + \Cw \sin(\omg t),
	\end{align*}
	where $\Cq, \Cw, \Dq, \Dw$ are constants, and we have defined
	\beq
		\omg = \frac{e B}{m c}.
	\eeq
	These solutions show that the particle moves in a circle with angular frequency $\omg$, as desired. \qed
\end{solution}

\begin{problem}
	Now consider the limit where the $B$ field can be made arbitrarily strong.  Compare the Poisson bracket $[x, \px]$ for the charged particle with the Poisson bracket relation
	\beq
		[\xi, \yi] = \frac{\delta_{ij}}{\gami}
	\eeq
	for the system of line vortices described in problems~\ref{vort1} and \ref{vort2}.
\end{problem}

\begin{solution}
	The Poisson bracket for the charged particle is given by
	\beq
		[x, \px] = \pdv{x}{x} \pdv{\px}{\px} - \pdv{x}{\px} \pdv{\px}{x} + \pdv{x}{y} \pdv{\px}{\py} - \pdv{x}{\py} \pdv{\px}{y}.
	\eeq
	Note that
	\begin{align*}
		\pdv{x}{x} &= \pdv{\px}{\px} = 1, &
		\pdv{x}{\px} &= \pdv{\px}{x} = \pdv{x}{y} = \pdv{\px}{\py} = 0, &
		\pdv{x}{\py} &= -\frac{c}{e B}, &
		\pdv{\px}{y} &= -\frac{e B}{c}.
	\end{align*}
	Then
	\beq
		[x, \px] = 1 - 1 = 0.
	\eeq
	\hl{It's different?}
\end{solution}


\vfill
While writing up these solutions, I consulted Gelfand and Fomin's \emph{Calculus of Variations}, Goldstein's \emph{Classical Mechanics}, and Riley, Hobson, and Bence's \emph{Mathematical Methods for Physics and Engineering}.
\end{document}