\documentclass[11pt]{article}
\usepackage{homework}

\classname{364}
\homeworknum{4}

\DeclareMathAlphabet{\mathsfit}{T1}{\sfdefault}{\mddefault}{\sldefault}


\begin{document}

\input{definitions}
\input{probdefs}

\state{}{
	Consider a particle which, as viewed by an observer in an inertial lab, is in a circular orbit in the $(x, y)$ plane with angular velocity $\omg$ and radius $r$.  Suppose that this particle carries a spin angular momentum $\vS$ (treated classically in this problem) which is Fermi-Walker transported.  Compute the time dependence of this angular momentum $\vSt$ where $t$ is the inertial time in the laboratory frame.  Show that in the non-relativistic limit, the complex vector $\Sx + i \Sy$ precesses about the $z$ axis with frequency $\omgT = r^2 \omg^3 / 2$.
}

\sol{
	Since the particle is in a circular orbit in the $xy$ plane, we can write its position in the lab frame as
	\eq{
		\vx = ( t, r \cos(\omg t), r \sin(\omg t), 0 ).
	}
	Then from MCP~(2.7), its velocity is
	\eqn{U}{
		\vU = \dv{\vx}{\tau}
		= \gam ( 1, -r \omg \sin(\omg t), r \omg \cos(\omg t), 0 ),
	}
	since $\ddtau = \ddt / \gam$~\cite[p.~201]{Resnick}.  The particle's acceleration is
	\eq{
		\vaa = \dv{\vU}{\tau}
		= -r \omg^2 \gam^2 ( 0, \cos(\omg t), \sin(\omg t), 0 ).
	}
	The Fermi-Walker transport is, as given by MCP~(24.62),
	\eqn{FW}{
		\nabsU \vS = \vU (\vaa \cdot \vS),
	}
	and we know that the spin vector is always orthogonal to the particle's 4-velocity~\cite[p.~1184]{MCP}.  This means that the spin vector can be written
	\eq{
		\vS = (0, \bS)
	}
	since $\vU \cdot \vS = 0$ is Lorentz invariant, and the spatial components of the particle's velocity are zero in its rest frame.  Moreover, this means
	\eq{
		\nabsU \vS = -\dv{\vS}{\tau}
		= -\gam \dv{\vS}{t},
	}
	so Eq.~\refeq{FW} can be written
	\eq{
		\dv{\vS}{t} = -\frac{1}{\gam} \vU (\vaa \cdot \vS)
		= \gam r \omg^2 [ \Sx \cos(\omg t) + \Sy \sin(\omg t) ] \vU.
	}
	Feeding in the relevant components of Eq.~\refeq{U}, we have the system of coupled differential equations
	\al{
		\dv{\Sx}{t} &= -\gam^2 r^2 \omg^3 \sin(\omg t) [ \Sx \cos(\omg t) + \Sy \sin(\omg t) ], \\
		\dv{\Sy}{t} &= \gam^2 r^2 \omg^3 \cos(\omg t) [ \Sx \cos(\omg t) + \Sy \sin(\omg t) ],
	}
	or, in matrix form,
	\eqn{sysxy}{
		\dv{t} \mqty[ \Sx \\ \Sy ] = \gam^2 r^2 \omg^3 \mqty[
				-\cos(\omg t) \sin(\omg t) & -\sin[2](\omg t) \\
				\cos[2](\omg t) & \cos(\omg t) \sin(\omg t)
			] \mqty[ \Sx \\ \Sy ].
	}
	To solve the system, we define the polar components of $\vS$ by~\cite[inside cover]{Griffiths}
	\al{
		\Sr &= \Sx \cos(\omg t) + \Sy \sin(\omg t), &
		\Stht &= -\Sx \sin(\omg t) + \Sy \cos(\omg t).
	}
	Then we can transform into these polar coordinates using the matrix~\cite[inside cover]{Griffiths}
	\eqn{sub1}{
		\mqty[ \Sx \\ \Sy ] = \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \mqty[ \Sr \\ \Stht ],
	}
	which implies
	\aln{
		\dv{t} \mqty[ \Sx \\ \Sy ] &= \dv{t} \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \mqty[ \Sr \\ \Stht ] + \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \dv{t} \mqty[ \Sr \\ \Stht ] \\ \notag
		&= -\omg \mqty[
			\sin(\omg t) & \cos(\omg t) \\
			-\cos(\omg t) & \sin(\omg t)
		] \mqty[ \Sr \\ \Stht ] + \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \dv{t} \mqty[ \Sr \\ \Stht ]. \label{sub2}
	}
	Substituting Eqs.~\refeq{sub1} and \refeq{sub2} into Eq.~\refeq{sysxy} yields
	\al{
		-\omg &\mqty[
				\sin(\omg t) & \cos(\omg t) \\
				-\cos(\omg t) & \sin(\omg t)
			] \mqty[ \Sr \\ \Stht ] + \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \dv{t} \mqty[ \Sr \\ \Stht ] \\
		&\hspace{10em} = \gam^2 r^2 \omg^3 \mqty[
				-\cos(\omg t) \sin(\omg t) & -\sin[2](\omg t) \\
				\cos[2](\omg t) & \cos(\omg t) \sin(\omg t)
			] \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \mqty[ \Sr \\ \Stht ] \\
		&\hspace{10em} = \gam^2 r^2 \omg^3 \mqty[
				-\sin(\omg t) & 0 \\
				\cos(\omg t) & 0
			] \mqty[ \Sr \\ \Stht ],
	}
	where the matrix multiplication has been carried out with Mathematica.  This implies
	\al{
		\mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \dv{t} \mqty[ \Sr \\ \Stht ]
		&= \paren{ \omg \mqty[
				\sin(\omg t) & \cos(\omg t) \\
				-\cos(\omg t) & \sin(\omg t)
			] + \gam^2 r^2 \omg^3 \mqty[
				-\sin(\omg t) & 0 \\
				\cos(\omg t) & 0
			] } \mqty[ \Sr \\ \Stht ] \\
		&= \omg \mqty[
				(1 - \gam^2 r^2 \omg^2) \sin(\omg t) & \cos(\omg t) \\
				-(1 - \gam^2 r^2 \omg^2) \cos(\omg t) & \sin(\omg t)
			] \mqty[ \Sr \\ \Stht ].
	}
	Multiplying both sides by the inverse of the first matrix, we find
	\eq{
		\dv{t} \mqty[ \Sr \\ \Stht ] = \omg \mqty[
				\cos(\omg t) & \sin(\omg t) \\
				-\sin(\omg t) & \cos(\omg t)
			] \mqty[
				(1 - \gam^2 r^2 \omg^2) \sin(\omg t) & \cos(\omg t) \\
				-(1 - \gam^2 r^2 \omg^2) \cos(\omg t) & \sin(\omg t)
			] \mqty[ \Sr \\ \Stht ]
		= \omg \mqty[
				0 & 1 \\
				\gam^2 r^2 \omg^2 - 1 & 0
			] \mqty[ \Sr \\ \Stht ].
	}
	In other words, we have a system of coupled first-order ODEs:
	\aln{ \label{ODEs}
		\dv{\Sr}{t} &= \omg \Stht, &
		\dv{\Stht}{t} &= \omg (\gam^2 r^2 \omg^2 - 1) \Sr.
	}
%	However, we know that $\Sr(t) = S = \const$, since the magnitude of the spin should not change with time.  Let $\alp \equiv \omg^2 (\gam^2 r^2 \omg^2 - 1)$.  Then, imposing the initial condition $\Stht(0) = 0$,
%	\al{
%		\Sr(t) &= S, &
%		\Stht(t) &= \alp S t.
%	}
%	Now we use Eq.~\refeq{sub1} to transform back to Cartesian components:
%	\al{
%		\Sx(t) &= S \cos(\omg t) - \alp S \sin(\omg t) t
%	}
%	
%	\hl{actually this doesn't make sense}
%
	Differentiating each equation by $t$ once more and substituting, we get a system of uncoupled second-order ODEs with well-known solutions:
	\al{
		\dv[2]{\Sr}{t} &= \omg \dv{\Stht}{t}
		= \omg^2 (\gam^2 r^2 \omg^2 - 1) \Sr, &
		%
		\dv[2]{\Stht}{t} &= \omg (\gam^2 r^2 \omg^2 - 1) \dv{\Sr}{t}
		= \omg^2 (\gam^2 r^2 \omg^2 - 1) \Stht.
	}
	Let $\alp \equiv \omg^2 (\gam^2 r^2 \omg^2 - 1)$.  Then the solutions are~\cite[p.~207]{Swartz}
	\al{
		\Sr(t) &= \Cq \cos(-\alp t) + \Cw \sin(-\alp t), &
		\Stht(t) &= \Dq \cos(-\alp t) + \Dw \sin(-\alp t).
	}
	We transform back into Cartesian components via Eq.~\refeq{sub1}:
	\al{
		\mqty[ \Sx \\ \Sy ] &= \mqty[
				\cos(\omg t) & -\sin(\omg t) \\
				\sin(\omg t) & \cos(\omg t)
			] \mqty[ 
				\Cq \cos(-\alp t) + \Cw \sin(-\alp t) \\
				\Dq \cos(-\alp t) + \Dw \sin(-\alp t)
			] \\
		&= \mqty[
				\cos(\omg t) [ \Cq \cos(-\alp t) + \Cw \sin(-\alp t) ] - \sin(\omg t) [ \Dq \cos(-\alp t) + \Dw \sin(-\alp t) ] \\
				\sin(\omg t) [ \Cq \cos(-\alp t) + \Cw \sin(-\alp t) ] + \cos(\omg t) [ \Dq \cos(-\alp t) + \Dw \sin(-\alp t) ].
			]
	}
	We choose the initial conditions $\Sx(0) = S$, $\Sy(0) = 0$, which give us
	\al{
		\Sx(0) &= S
		= \Cq, &
		%
		\Sy(0) &= 0
		= \Dq.
	}
	Using these results and the initial conditions in Eq.~\refeq{sysxy}, we have
	\eq{
		\left. \dv{t} \mqty[ \Sx \\ \Sy ] \right|_{t = 0} = \gam^2 r^2 \omg^3 \mqty[
				0 & 0 \\
				1 & 0
			] \mqty[
				S \\
				0
			]
		= \gam^2 r^2 \omg^3 \mqty[ 0 \\ S ],
	}
	with
	\eq{
		\left. \dv{t} \mqty[ \Sx \\ \Sy ] \right|_{t = 0} = 
	}
	\hl{but this is getting really gross}
	
	\eq{
		\mqty[ \Sx \\ \Sy ] = \mqty[
				\cos(\omg t) [ S \cos(-\alp t) + \Cw \sin(-\alp t) ] - \Dw \sin(\omg t) \sin(-\alp t) \\
				\sin(\omg t) [ S \cos(-\alp t)  + \Cw \sin(-\alp t) ] + \Dw \cos(\omg t) \sin(-\alp t)
			]
	}
	\eq{
		\mqty[ 0 \\ S ](t = -\pi \alp / 2) = \mqty[
				\Cw \cos(-\pi \omg \alp / 2) - \Dw \sin(-\pi \omg \alp / 2) \\
				\Cw \sin(-\pi \omg \alp / 2) + \Dw \cos(-\pi \omg \alp / 2)
			]
	}
	
	
%	We choose the initial conditions $\Sr(0) = S$ and $\Stht(0) = 0$.  This gives us
%	\al{
%		S &= \Sr(0) = \Cq, &
%		0 &= \Stht(0) = \Dq.
%	}
%	Then by Eq.~\refeq{ODEs},
%	\al{
%		\omg \Dw \sin(-\alp t) &= \dv{\Sr}{t} = \omg \alp [ S \sin(-\alp t) - \Cw \cos(-\alp t) ], \\
%		\omg \alp [ S \cos(-\alp t) + \Cw \sin(-\alp t) ] &= \dv{\Stht}{t} = -\alp \Dw \cos(-\alp t)
%	}
}





\clearpage
\state{Gravitational redshift~(MCP 24.16)}{
	Inside a laboratory on Earth's surface the effects of spacetime curvature are so small that current technology cannot measure them.  Therefore, experiments performed in the laboratory can be analyzed using special relativity. %  (This fact is embodied in Einstein's equivalence principle.)
}

\prob{
	Explain why the spacetime metric in the proper reference frame of the laboratory's floor has the form
	\eqn{metric2}{
		\dds^2 = (1 + 2 g z) (\ddxoh)^2 + \ddx^2 + \ddy^2 + \ddz^2,
	}
	plus terms due to the slow rotation of the laboratory walls, which we neglect in this exercise.  Here $g$ is the acceleration of gravity measured on the floor.
}

\sol{
	We can transform coordinates from the proper reference frame of the laboratory floor to another inertial frame.  We choose this other frame such that it is only a very small ``distance'' away at $\vx = 0$ in the proper frame.  That is, the frames are identical in the immediate vicinity of event (small $\vx$).  Then the coordinate transformation from the proper reference frame to the other inertial frame is given by MCP~(24.60a),
	\al{
		\xii &= \xiih + \frac{1}{2} \aih (\xoh)^2 + \epsisjkh \Omgjh \xkh \xoh, &
		\xo &= \xoh (1 + \asjh \xjh),
	}
	where terms to quadratic order in $\xah$ are included, and $\Omgjh$ is the rotational angular velocity of the laboratory.  Since the metric in the inertial frame is $\dds^2 = -(\ddxo)^2 + \delsij \ddxii \ddxj$~\cite[p.~1183]{MCP}, the metric in the proper reference frame is given by MCP~(24.60b),
	\eq{
		\dds^2 = -(1 + 2 \ba \vdot \bx) (\ddxoh)^2 + 2 (\bOmg \cross \bx) \vdot \ddbx \ddxoh + \delsjk \ddxjh \ddxkh,
	}
	which is accurate to linear order in $\xah$.  For this problem, we ignore rotations so $\bOmg \to \bo$, eliminating the second term.  Also, $\ba = g \zh$ so $\ba \vdot \bx = -g z$.  Finally, we note that $\xiih \in \{ x, y, z \}$ since these coordinates coincide with the inertial frame~\cite[p.~1186]{MCP}.  Then the spacetime metric is
	\eq{
		\dds^2 = -(1 + 2 g z) (\ddxoh)^2 + \ddx^2 + \ddy^2 + \ddz^2,
	}
	which is what we want.  Evidently, there is a typo (missing minus sign) in the problem statement. \qed
}


\prob{
	An electromagnetic wave is emitted from the floor, where it is measured to have wavelength $\lamo$, and is received at the ceiling.  Using the metric of Eq.~\refeq{metric2}, show that, as measured in the proper reference frame of an observer on the ceiling, the received wave has wavelength $\lamr = \lamo (1 + g h)$, where $h$ is the height of the ceiling above the floor (i.e., the light is \emph{gravitationally redshifted} by $\Del\lam / \lamo = g h$).
	
	Hint: Show that all crests of the wave must travel along world lines that have the same shape, $z = F(\xoh - \xohe)$, where $F$ is some function, and $\xohe$ is the coordinate time at which the crest is emitted from the floor. %  You can compute the shape function $F$ if you wish, but it is not needed to derive the gravitational redshift; only its universality is needed.]
}





\clearpage
\state{Rigidly rotating disk~(MCP 24.17)}{
	Consider a thin disk with radius $R$ at $z = 0$ in a Lorentz reference frame.  The disk rotates rigidly with angular velocity $\Omg$.  In the early years of special relativity there was much confusion over the geometry of the disk: In the inertial frame it has physical radius (proper distance from center to edge) $R$ and physical circumference $\sC = 2\pi R$.  But Lorentz contraction dictates that, as measured on the disk, the circumference should be $\sqrt{1 - v^2} \sC$ (with $v = \Omg R$), and the physical radius, $R$, should be unchanged.  This seemed weird.  How could an obviously flat disk in spacetime have a curved, non-Euclidean geometry, with physical circumference divided by physical radius smaller than $2\pi$?  In this exercise you will explore this issue.
}

\prob{
	Consider a family of observers who ride on the edge of the disk.  Construct a circular curve, orthogonal to their world lines, that travels around the disk (at $\sqrt{x^2 + y^2} = R$).  This curve can be thought of as lying in a 3-surface of constant time $\xoh$ of the observers' proper reference frames.  Show that it spirals upward in a Lorentz-frame spacetime diagram, so it cannot close on itself after traveling around the disk.  Thus the 3-planes, orthogonal to the observers' world lines at the edge of the disk, cannot mesh globally to form global 3-planes.
}

\sol{
	We can write the position of an observer in an inertial reference frame as
	\eq{
		\vx = ( t, R \cos(\Omg t), R \sin(\Omg t), 0 ).
	}
	Then the velocity of the observer, which is tangent to his/her worldline, is
	\eq{
		\vuu = \dv{\vx}{\tau}
		= \gam ( 1, -R \Omg \sin(\Omg t), R \Omg \cos(\Omg t), 0 ).
	}
	By inspection, a vector orthogonal to this velocity is
	\eq{
		\vv = \gam ( R^2 \Omg^2, -R \Omg \sin(\Omg t), R \Omg \cos(\Omg t), 0 )
		= \dv{\vy}{\tau},
	}
	where $\vy$ traces out the curve orthogonal to the world line.  It is given by
	\eq{
		\vy = ( R^2 \Omg^2 t , R \cos(\Omg t), R \sin(\Omg t), 0 ).
	}
	We note that $\vy$ traces out a helix in a Lorentz-frame spacetime diagram.  Thus, it cannot close on itself after traveling around the disk. \qed
}



\prob{
	Next, consider a 2-dimensional family of observers who ride on the surface of the rotating disk.  Show that at each radius $\sqrt{x^2 + y^2} = \const$, the constant-radius curve that is orthogonal to their world lines spirals upward in spacetime with a different slope.  Show that this means that even locally, the 3-planes orthogonal to each of their world lines cannot mesh to form larger 3-planes---thus there does not reside in spacetime any 3-surface orthogonal to these observers' world lines.  There is no 3-surface that has the claimed non-Euclidean geometry.
}

\sol{
	For a given radius $r = \sqrt{x^2 + y^2}$, the constant-radius curve that is orthogonal to the worldline is given by
	\eq{
		\vy = ( r^2 \Omg^2 t, -r \Omg \sin(\Omg t), r \Omg \cos(\Omg t), 0 )
	}
	The slope at which this curve spirals upward in the spacetime diagram is $r^2 \Omg^2 / r = r \Omg^2$, which has a linear dependence on the radius.  Thus, the slope is different for any radius $r$. \qed
	
	\hl{what about the meshing?}
}






\clearpage
\state{Constant of geodesic motion in a spacetime with symmetry~(MCP 25.4)}{\hfix}

\prob{
	Suppose that in some coordinate system the metric coefficients are independent of some specific coordinate $\xA$: $\sgsabA = 0$ (e.g., in spherical polar coordinates $\{ t, r, \tht, \phi \}$ in flat spacetime $\sgsabphi = 0$, so we could set $\xA = \phi$).  Show that
	\eq{
		\psA \equiv \vp \cdot \pdv{\xA}
	}
	is a constant of the motion for a freely moving particle [$\psp = (\text{conserved $z$~component of angular momentum})$] in the above, spherically symmetric example.]
	
	Hint: Show that the geodesic equation can be written in the form
	\eq{
		\dv{\psa}{\zet} - \Gamsman \pmu \pnu = 0,
	}
	where $\Gamsman$ is the covariant connection of Eqs.~(24.38c), (24.38d) with $\csabg = 0$, because we are using a coordinate basis.
%	
%	Note the analogy of the constant of motion $\psA$ with Hamiltonian mechanics; there, if the Hamiltonian is independent of $\xA$, then the generalized momentum $\psA$ is conserved; here, if the metric coefficients are independent of $\xA$, then the covariant component $\psA$ of the momentum is conserved.
}

\sol{
	The general form of the geodesic equation is given by MCP~(25.11c),
	\eq{
		\nabsvp \vp = 0.
	}
	Dotting both sides by $\vp$ yields
	\eq{
		0 = \vp \cdot \nabsvp \vp
		\qimplies
		0 = \pmu \nabsm \psa.
	}
	The covariant components of the gradient are given by (24.36), $A_{\alp; \bet} = A_{\alp, \bet} - \Gam^\mu{}_{\alp \bet} A_\mu$, where $A_{\alp, \bet} = \ptsb A_\alp$ and $A_{\alp; \bet} = \nabsb A_\alp$.  Applying this, we have
	\eqn{thing4a}{
		0 = \pmu (p_{\alp, \mu} - \Gamgsam \psg)
		= \pmu \ptsm \psa - \Gamgsam \pmu \psg.
	}
	Using $\vp / m = \dv*{\vx}{\tau}$ and $\tau = m \zeta$~\cite[p.~1202]{MCP}, we can rewrite the first term of Eq.~\refeq{thing4a}:
	\eq{
		\pmu \ptsm \psa = m \dv{\xm}{\tau} \ptsm \psa
		= m \dv{\psa}{\tau}
		= \dv{\psa}{\zet}.
	}
	For the second term of Eq.~\refeq{thing4a}, we multiply by the metric $\sgsba$ as in (24.38d), $\Gam^\mu{}_{\bet \gam} = \sg^{\mu \alp} \Gam_{\alp \bet \gam}$.  Then we have
	\eq{
		\Gamgsam \pmu \psg = \sggn \Gamsnam \pmu \psg
		= \Gamsnam \pmu \pnu
		= \Gamsman \pmu \pnu,
	}
	where in the final step we have relabeled indices.  Thus we can write Eq.~\refeq{thing4a} as
	\eqn{thing4a2}{
		\dv{\psa}{\zet} - \Gamsman \pmu \pnu = 0,
	}
	as recommended.
	
	Now we apply (24.38c); in a coordinate basis, it reduces to $\Gamsabg = (\sgsabg + \sgsagb - \sgsbga) / 2$.  Then for the second term of Eq.~\refeq{thing4a2}, we can write
	\al{
		\Gamsman \pmu \pnu &= \frac{1}{2} (\sg_{\mu \alp, \nu} + \sg_{\mu \nu, \alp} - \sg_{\alp \nu, \mu} ) \pmu \pnu \\
		&= \frac{1}{2} (\sg_{\nu \alp, \mu} + \sg_{\nu \mu, \alp} - \sg_{\alp \mu, \nu} ) \pmu \pnu \\
		&= \frac{1}{2} (\sg_{\alp \nu, \mu} + \sg_{\mu \nu, \alp} - \sg_{\mu \alp, \nu} ) \pmu \pnu \\
		&= \frac{1}{2} \sgsmna \pmu \pnu,
	}
	where we have again relabeled indices, and also used the symmetry of the metric.  Then for Eq.~\refeq{thing4a2}, we have
	\eq{
		\dv{\psa}{\zet} = \frac{1}{2} \sgsmna \pmu \pnu.
	}
	Therefore when $\alp = A$,
	\eq{
		\dv{\psA}{\zet} = \frac{1}{2} \sgsmnA \pmu \pnu = 0
		\qimplies
		\dv{\psA}{\tau} = 0
		\qimplies
		\psA = \const,
	}
	as we wanted to show~\cite[pp.~134--135]{Carroll}. \qed
}



\prob{
	As an example, consider a particle moving freely through a time-independent, Newtonian gravitational field.  In Ex.~25.18, we learn that such a gravitational field can be described in the language of general relativity by the spacetime metric
	\eq{
		\dds^2 = -(1 + 2 \Phi) \ddt^2 + (\delsjk + \hsjk) \ddxj \ddxk,
	}
	where $\Phi(x, y, z)$ is the time-independent Newtonian potential, and $\hsjk$ are contributions to the metric that are independent of the time coordinate $t$ and have magnitude of order $\abs{\Phi}$.  That the gravitational field is weak means $\abs{\Phi} \ll 1$ (or, in conventional units, $\abs{\Phi / c^2} \ll 1$).  The coordinates being used are Lorentz, aside from tiny corrections of order $\abs{\Phi}$, and as this exercise and Ex.~25.18 show, they coincide with the coordinates of the Newtonian theory of gravity.  Suppose that the particle has velocity $\vj \equiv \dd*{\xj}{t}$ through this coordinate system that is $\lesssim \abs{\Phi}^{1/2}$ and thus is small compared to the speed of light.  Because the metric is independent of the time coordinate $t$, the component $\pst$ of the particle's 4-momentum must be conserved along its world line.  Since throughout physics, the conserved quantity associated with time-translation invariance is always the energy, we expect that $\pst$, when evaluated accurate to first order in $\abs{\Phi}$, must be equal to the particle's conserved Newtonian energy, $E = m \Phi + m \vj \vk \delsjk / 2$, aside from some multiplicative and additive constants.  Show that this, indeed, is true, and evaluate the constants.
}





\clearpage
\state{Killing vector field~(MCP 25.5)}{
	A \emph{Killing vector field} is a coordinate-independent tool for exhibiting symmetries of the metric.  It is any vector field $\vxi$ that satisfies
	\eq{
		\xisascb + \xisbsca = 0
	}
	(i.e., any vector field whose symmetrized gradient vanishes).
}

\prob{ \label{5a}
	Let $\vxi$ be a vector field that might or might not be Killing.  Show, by construction, that it is possible to introduce a coordinate system in which $\vxi = \pdv*{\xA}$ for some coordinate $\xA$.
}

\sol{
	If we treat $\vxi$ as a singular vector, it must exist in a space tangent to some manifold at a point we call $\cP$.  We can imagine a projection of $\vxi$ onto the manifold; such a projection is a curve on the manifold.  We may choose the direction of this curve as a coordinate of the manifold and call it $\xA$.  The curve itself we call $\cP(\xA)$.  Then, by construction, $\vxi$ is tangent to $\cP(\xA)$.  This means that $\vxi$ must be the directional derivative along $\cP(\xA)$, which is $\pdv*{\xA}$.  Thus, $\vxi = \pdv*{\xA}$~\cite[pp.~1166--1167]{MCP}. \qed
}



\prob{
	Show that in the coordinate system of \ref{5a} the symmetrized gradient of $\vxi$ is $\xisascb + \xisbsca = \pdv{\sgsab}{\xA}$.  From this infer that a vector field $\vxi$ is Killing if and only if there exists a coordinate system in which (i) $\vxi = \pdv*{\xA}$ and (ii) the metric is independent of $\xA$.
}

\sol{
%	By MCP~(24.36),
%	\eq{
%		A_{\alp; \bet} = A_{\alp, \bet} - \Gammsab A_\mu.
%	}
%	So
%	\al{
%		\xisascb &= \xisacb - \Gammsab \xism
%		= \xisacb - \pdv{\GamAsab}{\xA}, &
%		\xisbsca &= \xisbca - \Gamnsba \xisn
%		= \xisbca - \pdv{\GamAsba}{\xA},
%	}
%	since $\vxi$ has a component in only the $\xA$ direction.  The basis of \ref{5a} is a coordinate basis, so the $c$ terms vanish in MCP~(24.38), which becomes
%	\eq{
%		\Gamsabg = \frac{1}{2} (\sgsabg + \sgsagb - \sgsbga),
%	}
%	and from (24.38d),
%	\eq{
%		\Gammsbg = \frac{1}{2} \sgma (\sgsabg + \sgsagb - \sgsbga)
%	}
	
	
%	the $\Gammsab$ are symmetric in their last two indices~\cite[p.~1172]{MCP}.  This means
%	\eq{
%		\xisascb + \xisbsca = \ptsvesb \xisa + \ptsvesa \xisb - 2 \Gammsab \xism
%	}
}



\prob{
	Use Killing's equation~(25.18) to show, without introducing a coordinate system, that, if $\vxi$ is a Killing vector field and $\vp$ is the 4-momentum of a freely-falling particle, then $\vxi \cdot \vp$ is conserved along the particle's geodesic world line.  This is the same conservation law as we proved in Ex.~25.4a using a coordinate-dependent calculation.
}


\makebib

\end{document}
