 \documentclass[11pt]{article}
\usepackage{geometry, titlesec}
\usepackage[parfill]{parskip}
\usepackage[italicdiff]{physics}
\usepackage{amsfonts, amsthm}
\usepackage[cm]{fullpage}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{xcolor, soul}
\usepackage{kbordermatrix}
\usepackage{booktabs}
%\allowdisplaybreaks

\makeatletter
\renewcommand*\env@cases[1][1.2]{%
  \let\@ifnextchar\new@ifnextchar
  \left\lbrace
  \def\arraystretch{#1}%
  \array{@{}l@{\quad}l@{}}%
}
\makeatother

 
\renewcommand{\footrulewidth}{.2pt}
%\setlist[enumerate]{leftmargin=*}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textbf{Physics 342 Homework 3}}
\rhead{Lacey Rainbolt}
\setlength{\headheight}{11pt}
\setlength{\headsep}{11pt}
\setlength{\footskip}{24pt}
\lfoot{\today}
\rfoot{\thepage}

\titleformat{\section}[runin]{\normalfont\large\bfseries}{Problem \thesection.}{1em}{}
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subparagraph}[leftmargin]{\normalfont\normalsize\bfseries}{}{0pt}{}

\newcommand{\refeq}[1]{(\ref{#1})}

\newcommand{\beq}{\begin{equation*}}
\newcommand{\eeq}{\end{equation*}}

\newcommand{\beqn}{\begin{equation}}
\newcommand{\eeqn}{\end{equation}}


\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vfix}{\vspace{-\baselineskip}}


\newenvironment{statement}[1]
{
	\section{#1}
	\color{darkgray}
	\ignorespaces
}
{
%    \smallskip
}

\newenvironment{problem}
{
	\color{darkgray}
	\subsection{}
	\ignorespaces
}


\newenvironment{solution}
{
	\paragraph{Solution.}
}
{
	\bigskip
}

\newcommand{\Schrodinger}{Schr\"{o}dinger}
\newcommand{\qimplies}{\quad \implies \quad}


\begin{document}

\input{Problem1}



\newcommand{\Vo}{V_0}
\newcommand{\vr}{\vb{r}}
\newcommand{\Rl}{R_l}
\newcommand{\Ylm}{Y_{l m}}
\newcommand{\Rlr}{\Rl(r)}
\newcommand{\vep}{\varepsilon}
%\newcommand{\tht}{\theta}
\newcommand{\del}{\delta}
\newcommand{\dell}{\del_l}
\newcommand{\delo}{\del_0}
\newcommand{\Al}{A_l}
\newcommand{\absr}{\abs{r}}

\newcommand{\psiE}{\psi_E}
\newcommand{\uE}{u_E}
\newcommand{\eff}{\text{eff}}
\newcommand{\Veff}{V_\eff}
\newcommand{\El}{E_l}

\newcommand{\jl}{j_l}
\newcommand{\jo}{j_0}
\newcommand{\nl}{n_l}
\newcommand{\hlq}{h_l^{(1)}}
\newcommand{\hoq}{h_0^{(1)}}
\newcommand{\Ro}{R_0}
\newcommand{\constant}{\text{constant }}
\newcommand{\kap}{\kappa}
\newcommand{\avep}{\abs{\vep}}

\newcommand{\betal}{\beta_l}


\clearpage
\begin{statement}{}
	Consider a quantum particle with mass $m$ moving in the presence of the square well potential
	\beq
		V(r) = \begin{cases}
			-\Vo & r \leq a, \\
			0 & r > a.
		\end{cases}
	\eeq
	\vfix
\end{statement}

\begin{problem}
	Writing the wave function in polar coordinates as $\psi(\vr) = \Rlr \, \Ylm(\tht, \phi)$, write down the {\Schrodinger} equation obeyed by $\Rl$.
\end{problem}

\begin{solution}
	From (A.5.1) in Sakurai, the full {\Schrodinger} equation is
	\beq
		-\frac{\hbar^2}{2m} \left[ \frac{1}{r^2} \pdv{}{r} \left( r^2 \pdv{\psiE}{r} \right) + \frac{1}{r^2 \sin\tht} \pdv{}{\tht} \left( \sin\tht \pdv{\psiE}{\tht} \right) + \frac{1}{r^2 \sin^2\tht} \pdv[2]{\psiE}{\phi} \right] + V(r) \,\psiE = E \psiE,
	\eeq
	where the angular part of $\psiE$ satisfies (A.5.4),
	\beq
		-\left[ \frac{1}{\sin\tht} \pdv{}{\tht} \left( \sin\tht \pdv{}{\tht} \right) + \frac{1}{\sin^2\tht} \pdv[2]{}{\phi} \right] \Ylm = l (l + 1) \Ylm.
	\eeq
	Then the equivalent one-dimensional {\Schrodinger} equation is the equation immediately following (A.5.8),
	\beq
		-\frac{\hbar^2}{2m} \dv[2]{\uE}{r} + \left[ V(r) + \frac{l (l + 1) \hbar^2}{2 m r^2} \right] \uE = E \uE,
	\eeq
	where $\uE(r) = r R(r)$.  In terms of $\Rl$,
	\beq
		-\frac{\hbar^2}{2m} \dv[2]{r} (r \Rl) + \left[ V(r) + \frac{l (l + 1) \hbar^2}{2 m r^2} \right] r \Rl = E r \Rl.
	\eeq
	or
	\beq
		\frac{\hbar^2}{2 m} \left[ -\frac{1}{r^2} \pdv{r} \left( r^2 \pdv{r} \right) + V(r) + \frac{l (l + 1)}{r^2} \right] \Rl(r) = \El \, \Rl(r).
	\eeq
	From (7.7.1), the effective potential at low energies for the $l$th partial wave is
	\beq
		\Veff = V(r) + \frac{\hbar^2}{2m} \frac{l (l + 1)}{r^2},
	\eeq
	so the {\Schrodinger} equation can be rewritten as
	\beq
		\left[ -\frac{\hbar^2}{2 m} \frac{1}{r^2} \pdv{}{r} \left( r^2 \pdv{}{r} \right) + \Veff\, \right] \Rl(r) = \El \, \Rl(r).
	\eeq
	\vfix
\end{solution}

\begin{problem}
	When $\Vo$ is a certain value, there is one bound state for the $s$ wave ($l = 0$).  The bound state energy $\vep$ is small ($0 < \avep \ll \Vo$).  Obtain the range of the depth of the well $\Vo$ (? $\leq \Vo <$ ?).  Also, calculate for the bound state the probability for the particle to exist outside of the well.
\end{problem}

\begin{solution}
	Inside the well, $\Rl$ are given by (A.5.16),
	\beq
		\Rl(r) = \constant \jl(\alp r),
	\eeq
	where $\alp$ is defined in Eq.~(A.5.17),
	\beq
		\alp = \sqrt{\frac{2m (\Vo - \abs{E})}{\hbar^2}}, \quad r < a,
	\eeq
	and the spherical Bessel functions $\jl$ is given by (A.5.12),
	\begin{align*}
		\jl(\rho) = \sqrt{\frac{\pi}{2\rho}} J_{l + 1/2}(\rho).
	\end{align*}
	Outside the well, $\Rl$ are given by (A.5.21),
	\beq
		\Rl(r) = \constant \hlq(i \kap r) = \constant [ \jl(i \kap r) + i \nl(i \kap r)],
	\eeq
	where $\kap$ is defined in (A.5.20),
	\beq
		\kap = \sqrt{\frac{2 m \abs{E}}{\hbar^2}},
	\eeq
	and $\hlq$ are the spherical Hankel functions.
	
	We need to match the inside and outside solutions at the boundary $r = a$ and determine the range of $\Vo$ for which exactly one solution exists.  For the $s$ wave, the relevant Bessel function is given by (A.5.12),
	\beq
		\jo(\rho) = \frac{\sin\rho}{\rho}.
	\eeq
	Then we have
	\beqn \label{inside}
		\Ro(r) \propto \sqrt{\frac{\hbar^2}{2 m (\Vo - \avep)}} \frac{1}{r} \sin(\sqrt{\frac{2 m \Vo}{\hbar^2}} r), \quad r < a,
	\eeqn
	for the solution inside the well.  For the solution outside, the relevant Hankel function is given by (A.5.22),
	\beq
		\hoq(i \kap r) = -\frac{1}{\kap r} e^{-\kap r}.
	\eeq
	Then we have
	\beqn \label{outside}
		\Ro(r) \propto -\sqrt{\frac{\hbar^2}{2 m \avep}} \frac{1}{r} \exp(-\sqrt{\frac{2 m \avep}{\hbar^2}} r), \quad r > a.
	\eeqn
	Matching \refeq{inside} and \refeq{outside} at $r = a$, we obtain
	\beq
		\sqrt{\frac{\hbar^2}{2 m \Vo}} \frac{1}{a} \sin(\sqrt{\frac{2 m (\Vo - \avep)}{\hbar^2}} a) = -\sqrt{\frac{\hbar^2}{2 m \avep}} \frac{1}{a} \exp(-\sqrt{\frac{2 m \avep}{\hbar^2}} a),
	\eeq
	which simplifies to
	\beq
		\sqrt{\frac{\avep}{\Vo - \avep}} \sin(\sqrt{\frac{2 m \Vo}{\hbar^2}} a) = -\exp(-\sqrt{\frac{2 m \avep}{\hbar^2}} a).
	\eeq
	Since $\avep \ll \Vo$, we may Taylor expand both sides about $\avep$.  To first order, we obtain
	\beq
		\sqrt{\frac{\avep}{\Vo}} \sin(\sqrt{\frac{2 m \Vo}{\hbar^2}} a) = -1 + \sqrt{\frac{2 m \avep}{\hbar^2}} a
	\eeq
	\hl{help???}
\end{solution}



\begin{problem}
	Consider the scattering problem by the well.  For each $l$, for large enough $r$, when $\Rlr$ is given by $\Rlr \sim \Al \sin(kr - l\pi / 2 + \dell) / r$, $\dell$ is called the scattering phase shift.  For the value of $\Vo$ within the range you obtained in the above problem, when the energy of the incident wave is is $E = 9 \Vo / 16$, calculate $\tan\delo$~(where $\delo$ is the scattering phase shift for the $s$ wave).
\end{problem}

\begin{solution}
	From (7.6.35) in Sakurai,
	\beq
		\tan\dell = \frac{k R \jl'(kR) - \betal \jl(kR)}{k R \nl'(kR) - \betal \nl(kR)},
	\eeq
	where (7.6.34) defines $\betal$,
	\beq
		\betal \equiv \left[ \frac{r}{\Al} \dv{\Al}{r} \right]_{r=R},
	\eeq
	and
	\beq
		\jl'(kR) = \left. \dv{\jl}{(kr)} \right|_{kr = kR}
	\eeq
\end{solution}

\begin{problem}
	Now consider the $S$ matrix, $S \equiv \exp(2i \delo) = \exp(i \delo) / \exp(-i \delo)$.  Compare the condition on $s$ wave bound state energies and the zero of the denominator of $S$.  Explain their relation.
\end{problem}




\newcommand{\gam}{\gamma}
\newcommand{\chio}{\chi_0}
\newcommand{\chior}{\chio(r)}
\newcommand{\Gam}{\Gamma}

\begin{statement}{}
	Consider a three dimensional potential
	\beq
		V(\absr) = \frac{\hbar^2 \gam}{2m} \del(\absr - a).
	\eeq
	The $s$ wave {\Schrodinger} equation is given by
	\beq
		-\frac{\hbar^2}{2m} \dv[2]{\chior}{r} + \frac{\hbar^2 \gam}{2m} \del(r - a) \, \chior = E \,\chior.
	\eeq
	The $s$ wave function must be regular (zero) at $r = 0$.  At $r = a$, it is continuous, but its derivative can jump.
\end{statement}

\begin{problem}
	Calculate the $s$ wave scattering phase shift $\do(k)$, where $k$ is related to $E$ as $E = \hbar^2 k^2 / 2m$.
\end{problem}

\begin{problem}
	When $\gam \gg k$, $1/a$ and when $\sin ka$ is not small, discuss the behavior of the scattering phase shift.
\end{problem}

\begin{problem}
	Obtain the condition to have resonant states and calculate the energy of the resonant states.
\end{problem}

\begin{problem}
	Calculate the width $\Gam$ of the resonance.  Discuss its behavior when $\gam$ is big.
\end{problem}

\begin{problem}
	When the velocity of the incident wave is small, obtain the scattering cross section.
\end{problem}



\vfill
I consulted Sakurai's \emph{Modern Quantum Mechanics}, Shankar's \emph{Principles of Quantum Mechanics}, and the Wikipedia article on a particle in a spherically symmetric potential while writing up these solutions.

\end{document}