\documentclass[11pt]{article}
\usepackage{geometry, titlesec}
\usepackage[parfill]{parskip}
\usepackage[italicdiff]{physics}
\usepackage{amsfonts, amsthm, mathrsfs}
\usepackage[cm]{fullpage}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{xcolor, soul}
\usepackage{siunitx}
\allowdisplaybreaks

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vfix}{\vspace{-\baselineskip}}

\makeatletter
\renewcommand*\env@cases[1][1.2]{%
  \let\@ifnextchar\new@ifnextchar
  \left\lbrace
  \def\arraystretch{#1}%
  \array{@{}l@{\quad}l@{}}%
}
\makeatother
 
\renewcommand{\footrulewidth}{.2pt}
%\setlist[enumerate]{leftmargin=*}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textbf{Physics 322 Homework 2}}
\rhead{Lacey Rainbolt}
\setlength{\headheight}{11pt}
\setlength{\headsep}{11pt}
\setlength{\footskip}{24pt}
\lfoot{\today}
\rfoot{\thepage}

\titleformat{\section}[runin]{\normalfont\large\bfseries}{Problem \thesection.}{1em}{}
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subparagraph}[leftmargin]{\normalfont\normalsize\bfseries}{}{0pt}{}

\newcommand{\refeq}[1]{(\ref{#1})}

\newcommand{\beq}{\begin{equation*}}
\newcommand{\eeq}{\end{equation*}}

\newcommand{\beqn}{\begin{equation}}
\newcommand{\eeqn}{\end{equation}}

\newcommand{\blg}{\begin{align*}}
\newcommand{\elg}{\end{align*}}


\newenvironment{statement}[1]
{
	\section{#1}
	\color{darkgray}
	\ignorespaces
}
{
%    \smallskip
}

\newenvironment{problem}
{
	\subsection{}
	\color{darkgray}
%	\paragraph{Problem.}
    \ignorespaces
}
{

}

\newenvironment{solution}
{
    \paragraph{Solution.}
    \ignorespaces
}
{
    \bigskip
}


\DeclareSIUnit\esu{esu}
\DeclareSIUnit\erg{erg}

\begin{document}


\input{Problem1}


\newcommand{\cV}{\mathcal{V}}
\newcommand{\phio}{\phi_0}
\newcommand{\phiox}{\phio(\vx)}
\newcommand{\const}{\text{const.}}
\newcommand{\sig}{\sigma}
\newcommand{\alp}{\alpha}
\newcommand{\sigtv}{\sig(\tht, \vph)}

\newcommand{\sE}{\mathscr{E}}
\newcommand{\dcx}{\dd[3]{x}}
\newcommand{\rhoo}{\rho_0}
\newcommand{\dS}{\dd{S}}
\newcommand{\evS}{|_S}
\newcommand{\evV}{|_\cV}
\newcommand{\intS}{\int_S}
\newcommand{\intV}{\int_\cV}

\begin{statement}{}
	Let $\cV$ be an arbitrary bounded region of space and suppose that a total charge $Q$ is to be distributed in $\cV$ in an arbitrary way, with $\rho = 0$ outside of $\cV$.  Show that the total energy is minimized if the charge is distributed the way that it would be if $\cV$ were a conductor, so that $\phi = \const$ within $\cV$ (and thus, in particular, all of the charge lies on the boundary of $\cV$).
	
	Hint: Let $\phiox$ be the potential one would obtain if $\cV$ were filled by a conducting body.  Consider the energy of $\phio + \phi'$, where the source $\rho'$ of $\phi'$ vanishes outside of $\cV$ and has no net charge within $\cV$.
\end{statement}

\begin{solution}
	Let $S = \partial \cV$ denote the boundary of $\cV$.  We separate space into three mutually exclusive regions: $\cV$, $S$, and the region outside (in which we are not interested).  By the superposition principle, we may write
	\begin{align*}
		\rho &= \rhoo + \rho', &
		\phi &= \phio + \phi',
	\end{align*}
	where $\rhoo$ is the charge of a conducting body filling $\cV$, $\phio$ is the electrostatic potential due to $\rhoo$, $\rho'$ is the charge distribution within $\cV$, and $\phi'$ is the electrostatic potential due to $\rho'$.  In order to eliminate ambiguity on the boundary, we require
	\begin{align} \label{rhoconds}
		\rhoo\evV &= 0, &
		\rho'\evS &= 0.
	\end{align}
	That is, $\rhoo = 0$ inside the conductor by definition, and $\rho'$ vanishes on the boundary where $\rhoo$ is nonzero.  For the entire body to have charge $Q$, we need
	\beq
		\int \rhoo \dcx = \intV \rho' \dcx + \intS \rhoo \dS = Q.
	\eeq
	From \refeq{rhoconds}, it follows that
	\beq
		\phio\evV = \phio\evS = \const
	\eeq

	The total energy is given by Eq.~(2.25) in the course notes,
	\beq
		\sE = \frac{1}{2} \int \phi \rho \dcx.
	\eeq
	So
	\begin{align*}
		\sE &= \frac{1}{2} \int (\phio + \phi') (\rhoo + \rho') \dcx
		= \frac{1}{2} \left( \int \phio (\rhoo + \rho') \dcx + \int \phi' (\rhoo + \rho') \dcx \right) \\
		&= \frac{1}{2} \left( \phio Q + \intV \phi' \rho' \dcx + \intS \phi' \rhoo \dS \right).
	\end{align*}
	\hl{help}
\end{solution}



\newcommand{\lap}{\nabla^2}
\newcommand{\Gd}{G_D}
\newcommand{\Gdxxp}{\Gd(\vx, \vx')}
\newcommand{\Gdxpx}{\Gd(\vx', \vx)}
\newcommand{\psixp}{\psi(\vx')}
\newcommand{\nh}{\vec{\hat{n}}}
\newcommand{\nabxp}{\nabla_{x'}}
\newcommand{\dSxp}{\dd{S_{x'}}}
\newcommand{\absxp}{\abs{\vx'}}

\begin{statement}{}
	Charge is distributed on a (nonconducting) sphere of radius $R$, i.e., the charge density throughout space is of the form $\rhox = \sig(\tht, \vph) \, \delta(r - R)$.  The surface charge distribution $\sig$ on the sphere is chosen in such a way that the electrostatic potential on the sphere is $\phi(r=R, \tht, \vph) = \alp \cost$, where $\alp$ is a constant.
\end{statement}

\begin{problem}
	Find the electrostatic potential $\phix$ at all $r \leq R$.
\end{problem}

\begin{solution}
	This is a Dirichlet boundary value problem.  We are seeking the solution to Poisson's equation $\lap\phi = -4\pi\rho$ subject to $\phi\evS = \psi = \alp \cost$.  Equation~(2.100) in the lecture notes gives the general solution,
	\beq
		\phix = \intV \Gdxxp \, \rhoxp \dcxp - \frac{1}{4\pi} \intS \psixp \, \nh' \cdot \nabxp \Gdxpx \dSxp.
	\eeq
	The Green's function for a sphere is given by Eq.~(2.91),
	\beq
		\Gdxxp = \frac{1}{\abs{\vx - \vx'}} + \frac{\beta}{\abs{\vx - \vx''}} \qq{where} \vx'' = \vx' \frac{R^2}{\absxp^2} \qand \beta = - \frac{R}{\absxp}.
	\eeq
	\hl{Spherical harmonic expansion (2.78)?}
\end{solution}


\begin{problem}
	Find the electrostatic potential $\phix$ at all $r \geq R$.
\end{problem}

\begin{problem}
	Find the surface charge density $\sigtv$ that was required in order to produce this potential $\phi$.
\end{problem}

\begin{problem}
	Find the total electrostatic energy.
\end{problem}



\newcommand{\vF}{\vec{F}}
\newcommand{\vE}{\vec{E}}
\newcommand{\rh}{\vec{\hat{r}}}
\newcommand{\thh}{\boldsymbol{\hat{\tht}}}
\newcommand{\phh}{\boldsymbol{\hat{\vph}}}
\newcommand{\nabx}{\nabla_{\!x}}
\newcommand{\Er}{E_r}

\begin{statement}{}
	A point charge of charge $q$ is placed at point $\vx'$ inside a conducting spherical shell of radius $R$.  There is no net charge on the conductor.  The potential inside the sphere is thus given by $q \, \Gdxxp$, where the explicit formula for $\Gdxxp$ for a spherical cavity is given in the lecture notes.
\end{statement}

\begin{problem}
	Find the surface charge density $\sigtv$ on the conducting shell.
\end{problem}

\begin{solution}
	The Green's function for a spherical cavity is given by Eq.~(2.91),
	\beq
		\Gdxxp = \frac{1}{\abs{\vx - \vx'}} + \frac{\alp}{\abs{\vx - \vx''}} \qq{where} \vx'' = \vx' \frac{R^2}{\absxp^2} \qand \alp = - \frac{R}{\absxp}.
	\eeq
	The surface charge density can be found from Eq.~(2.86),
	\beqn \label{scdeq}
		\vE \cdot \nh = 4\pi \sig,
	\eeqn
	where $\vE = -\nabla \phi$ in electrostatics.
	
	We will begin by finding $\vE$.  We will orient our coordinate system such that $\vx'$ (and consequently $\vx''$) points along the $z$ axis.  Note that
	\beq
		\Gdxxp = \frac{1}{\abs{\vx - \vx'}} - \frac{R}{\absxp \abs{\vx - \dfrac{R^2}{\absxp^2} \vx'}}
		= \frac{1}{\sqrt{\vx^2 - 2 \vx \cdot \vx' + {\vx'}^2}} - \frac{R}{\absxp \sqrt{\vx^2 - 2 \dfrac{R^2}{{\vx'}^2} \vx \cdot \vx' + \dfrac{R^4}{{\vx'}^4} {\vx'}^2}}.
	\eeq
	In spherical coordinates, we have
	\beq
		\Gdxxp = \frac{1}{\sqrt{r^2 - 2 r r' \cost + {r'}^2}} - \frac{R}{r'} \frac{1}{\sqrt{r^2 - 2 R^2 r \cost / r' + R^4 / {r'}^2}},
	\eeq
	where we note that $\tht$ is the angle between $\vx$ and the $z$ axis.  The gradient in spherical coordinates is given by
	\beq
		\nabla = \pdv{}{r} \,\rh + \frac{1}{r} \pdv{}{\tht} \,\thh + \frac{1}{r \sint} \pdv{}{\vph} \, \phh.
	\eeq
	The $r$ component of the electric field inside the conductor is then
	\beq
		\Er(\vx) = -q \pdv{\Gdxxp}{r}
		= q \left( \frac{r - r' \cost}{(r^2 - 2 r r' \cost + {r'}^2)^{3/2}} - \frac{R}{r'} \frac{r - R^2 \cost / r'}{(r^2 - 2 R^2 r \cost / r' + R^4 / {r'}^2)^{3/2}} \right).
	\eeq
	Since $\nh = -\rh$ for the inner surface of a sphere, we are interested in only the $r$ component of the field.  On the surface of the sphere, the field is $\Er(r=R) \,\rh$.  So we have
	\begin{align*}
		\Er(r=R) &= q \left( \frac{R - r' \cost}{(R^2 - 2 R r' \cost + {r'}^2)^{3/2}} - \frac{R}{r'} \frac{R - R^2 \cost / r'}{(R^2 - 2 R^3 \cost / r' + R^4 / {r'}^2)^{3/2}} \right) \\
		&= q \left( \frac{R - r' \cost}{{r'}^3 (R^2 / {r'}^2 - 2 R \cost / r' + 1)^{3/2}} - \frac{R}{r'} \frac{R - R^2 \cost / r'}{R^3 (1 - 2 R \cost / r' + R^2 / {r'}^2)^{3/2}} \right) \\
		&= \frac{q}{r'} \frac{R^3 - R^2 r' \cost - R {r'}^2 + R^2 r' \cost}{R^2 {r'}^2 (R^2 / {r'}^2 - 2 R \cost / r' + 1)^{3/2}}
		= \frac{q}{R {r'}^3} \frac{R^2 - {r'}^2}{(R^2 / {r'}^2 - 2 R \cost / r' + 1)^{3/2}}.
	\end{align*}
	Finally, feeding this into \refeq{scdeq},
	\beq
		\sig = -\frac{\vE \cdot \rh}{4\pi}
		= \frac{q}{4\pi R {r'}^3} \frac{{r'}^2 - R^2}{(R^2 / {r'}^2 - 2 R \cost / r' + 1)^{3/2}}
		= \frac{q}{4\pi R \absxp^3} \frac{\absxp^2 - R^2}{(R^2 / \absxp^2 - 2 R \cost / \absxp + 1)^{3/2}}.
	\eeq
\end{solution}
\vfix


\newcommand{\vEo}{\vE_0}
\newcommand{\del}{\delta}
\newcommand{\Etht}{E_\tht}
\newcommand{\Fr}{F_r}

\begin{problem}
	Find the force $\vF$ that must be exerted on the point charge in order to hold it in place.
\end{problem}

\begin{solution}
	The total force on a charge distribution arises only from the external electric field $\vEo$, and is given by Eq.~(2.42) in the lecture notes:
	\beq
		\vF = \int \rhox \, \vEo(\vx) \dcx.
	\eeq
	We now need the $\tht$ component of the field inside the conductor, which is
	\beq
		\Etht(\vx) = -\frac{q}{r} \pdv{\Gdxxp}{\tht}
		= -q \left( \frac{r' \sint}{(r^2 - 2 r r' \cost + {r'}^2)^{3/2}} - \frac{R^3 \sint}{{r'}^2 (r^2 - 2 R^2 r \cost / r' + R^4 / {r'}^2)^{3/2}} \right).
	\eeq
	The charge density for a point charge located at $\vx'$ is given by $\rhox = q \, \delta(\vx - \vx')$.  Evaluating the integral, we have
	\beq
		\vF = \int q \, \delta(\vx - \vx') \, \vE(\vx) \dcx = q \vE(\vx').
	\eeq
	Recall that we chose $\vx'$ to point along the $z$ axis, so $\tht' = 0$.  The $\tht$ component of $\vF$ is then $0$, and the $r$ component is
	\begin{align*}
		\Fr &= q^2 \left( \frac{r' - r'}{({r'}^2 - 2 {r'}^2 + {r'}^2)^{3/2}} - \frac{R}{r'} \frac{r' - R^2 / r'}{({r'}^2 - 2 R^2 + R^4 / {r'}^2)^{3/2}} \right)
		= q^2 R {r'}^2 \frac{R^2 / r' - r'}{({r'}^4 - 2 R^2 {r'}^2 + R^4)^{3/2}} \\
		&= q^2 R {r'}^2 \frac{(R^2 - {r'}^2) / r'}{(R^2 - {r'}^2)^3}
		= q^2 \frac{R r'}{(R^2 - {r'}^2)^2}.
	\end{align*}
	Since only the $r$ component of $\vF$ is nonzero, it points in the $z$ direction, which we chose to be equivalent to the unit vector $\vx' / \absxp$.  Therefore,
	\beq
		\vF = q^2 \frac{R \absxp}{(R^2 - \absxp^2)^2} \frac{\vx'}{\absxp}
		= q^2 \frac{R}{(R^2 - \absxp^2)^2} \vx'.
	\eeq
\end{solution}




\begin{statement}{}
	The ``mean value theorem'' is stated as follows: For any solution $\phi$ to $\lap \phi = 0$, the value of $\phi$ at $\vx$ is equal to the average value of $\phi$ on a sphere of radius $R$ (for any $R$) centered at $\vx$.
\end{statement}

\begin{problem}
	Prove the mean value theorem.  Hint: Apply Green's theorem to $\phi$ and $1/\abs{\vx - \vx'}$ for a suitable choice of region and a suitable choice of $\vx'$.
\end{problem}

\begin{problem}
	Use this result to show that a point charge can never be in stable equilibrium if placed in an electric field $\vE$ that is source free in a neighborhood of the charge---and, indeed, it can be in neutral equilibrium only if $\vE = 0$ in this neighborhood.
\end{problem}

%\vfill
%In addition to the course lecture notes, I consulted Griffiths's \emph{Introduction to Electrodynamics} and David Tong's lecture notes on electrodynamics while writing up these solutions.
\end{document}