\documentclass[11pt]{article}
\usepackage{geometry, titlesec}
\usepackage[parfill]{parskip}
\usepackage[italicdiff]{physics}
\usepackage{amsfonts, amsthm, mathrsfs}
\usepackage[cm]{fullpage}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{xcolor, soul}
\usepackage{siunitx, graphicx}
%\allowdisplaybreaks

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\newcommand{\vfix}{\vspace{-\baselineskip}}
\newcommand{\qimplies}{\quad\implies\quad}

\makeatletter
\renewcommand*\env@cases[1][1.2]{%
  \let\@ifnextchar\new@ifnextchar
  \left\lbrace
  \def\arraystretch{#1}%
  \array{@{}l@{\quad}l@{}}%
}
\makeatother
 
\renewcommand{\footrulewidth}{.2pt}
%\setlist[enumerate]{leftmargin=*}
\pagestyle{fancy}
\fancyhf{}
\lhead{\textbf{Physics 323 Homework 1}}
\rhead{Lacey Rainbolt}
\setlength{\headheight}{11pt}
\setlength{\headsep}{11pt}
\setlength{\footskip}{24pt}
\lfoot{\today}
\rfoot{\thepage}

\titleformat{\section}[runin]{\normalfont\large\bfseries}{Problem \thesection}{1em}{}
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subparagraph}[leftmargin]{\normalfont\normalsize\bfseries}{}{0pt}{}

\newcommand{\refeq}[1]{(\ref{#1})}

\newcommand{\beq}{\begin{equation*}}
\newcommand{\eeq}{\end{equation*}}

\newcommand{\beqn}{\begin{equation}}
\newcommand{\eeqn}{\end{equation}}

\newcommand{\blg}{\begin{align*}}
\newcommand{\elg}{\end{align*}}


\newenvironment{statement}[1]
{
	\section{#1.}
	\color{darkgray}
	\ignorespaces
}

\newenvironment{problem}
{
	\subsection{}
	\color{darkgray}
	\ignorespaces
}

\newenvironment{solution}
{
	\paragraph{Solution.}
	\ignorespaces
}
{
    \bigskip
}

\DeclareSIUnit{\MeV}{\mega\electronvolt}
\DeclareSIUnit{\GeV}{\giga\electronvolt}

\begin{document}

\newcommand{\alp}{\alpha}
\newcommand{\bet}{\beta}
\newcommand{\gam}{\gamma}
\newcommand{\del}{\delta}
\newcommand{\lam}{\lambda}
\newcommand{\tht}{\theta}

\newcommand{\Del}{\Delta}
\newcommand{\Lam}{\Lambda}
\newcommand{\Tht}{\Theta}

\newcommand{\etamn}{\eta^{\mu \nu}}


\input{Problem1}

\input{Problem2}





\newcommand{\vr}{\vb{r}}
\newcommand{\vrperp}{\vr_\perp}

\begin{statement}{(Jackson 11.18)}
	The electric and magnetic fields of a particle of charge $q$ moving in a straight line with speed $v = \bet c$, given by \refeq{fields}, become more and more concentrated as $\bet \to 1$.  Choose axes so that the charge moves along the $z$ axis in the positive direction, passing the origin at $t = 0$.  Let the spatial coordinates of the observation point be $(x, y, z)$ and define the transverse vector $\vrperp$, with components $x$ and $y$.  Consider the fields and the source in the limit of $\bet = 1$.
\end{statement}

\newcommand{\rperp}{r_\perp}
\newcommand{\vE}{\vb{E}}
\newcommand{\vB}{\vb{B}}
\newcommand{\vh}{\vb{\hat{v}}}
\newcommand{\delctz}{\del(ct - z)}

\begin{problem} \label{3.a}
	Show that the fields can be written as
	\begin{align*}
		\vE &= 2q \frac{\vrperp}{\rperp^2} \delctz, &
		\vB &= 2q \frac{\vh \cross \vrperp}{\rperp^2} \delctz,
	\end{align*}
	where $\vh$ is a unit vector in the direction of the particle's velocity.
\end{problem}

\newcommand{\Ja}{J^\alp}
\newcommand{\Jb}{J^\bet}
\newcommand{\dela}{\partial_\alp}
\newcommand{\valp}{v^\alp}

\newcommand{\rperpq}{{\rperp}_1}
\newcommand{\rperpw}{{\rperp}_2}

\newcommand{\delt}{\partial_t}
\newcommand{\delx}{\partial_x}
\newcommand{\dely}{\partial_y}
\newcommand{\delz}{\partial_z}

\begin{problem}
	Show that by substitution into the Maxwell equations that these fields are consistent with a 4-vector source density,
	\beq
		\Ja = q c \valp \del^2(\vrperp) \delctz,
	\eeq
	where the 4-vector $\valp = (1, \vh)$.
\end{problem}

\begin{solution}
	According to Jackson~(11.141), the inhomogeneous Maxwell equations can be written
	\beq
		\dela \Fab = \frac{4\pi}{c} \Jb.
	\eeq
	Here,
	\beq
		\Fab %= \frac{2q}{{\rperp}^2} \delctz
%			\mqty[ 	0 & -\rperpq & -\rperpw & 0 \\
%					\rperpq & 0 & 0 & \rperpq \\
%					\rperpw & 0 & 0 & \rperpw \\
%					0 & -\rperpq & -\rperpw & 0 ]
		= \frac{2q}{{\rperp}^2} \delctz
			\mqty[ 	0 & -x & -y & 0 \\
					x & 0 & 0 & x \\
					y & 0 & 0 & y \\
					0 & -x & -y & 0 ]	
	\eeq
	and so
	\beq
		\dela \Fab = \mqty[ \delt & -\delx & -\dely & -\delz] \frac{2q}{{\rperp}^2} \delctz
			\mqty[ 	0 & -x & -y & 0 \\
					x & 0 & 0 & x \\
					y & 0 & 0 & y \\
					0 & -x & -y & 0 ]	
		= \mqty[ & & & ]
	\eeq
\end{solution}




\newcommand{\Ao}{A^0}
\newcommand{\Az}{A^z}
\newcommand{\vA}{\vb{A}}
\newcommand{\vAperp}{\vA_\perp}
\newcommand{\Thtctz}{\Tht(ct - z)}
\newcommand{\grperp}{\grad_\perp}

\begin{problem}
	Show that the fields of \refeq{3.a} are derivable from either of the following 4-vector potentials,
	\begin{align*}
		\Ao &= \Az = -2q \delctz \ln(\lam \rperp), &
		\vAperp &= 0,
	\end{align*}
	or
	\begin{align*}
		\Ao &= 0 = \Az, &
		\vAperp &= -2q \Thtctz \grperp \ln(\lam \rperp),
	\end{align*}
	where $\lam$ is an irrelevant parameter setting the scale of the logarithm.
	
	Show that the two potentials differ by a gauge transformation and find the gauge function, $\chi$.
\end{problem}




\clearpage

\newcommand{\mq}{m_1}
\newcommand{\mw}{m_2}
\newcommand{\pq}{p_1}
\newcommand{\pw}{p_2}

\newcommand{\mLam}{\SI{1115}{\MeV}}
\newcommand{\tauLam}{\SI{2.9e-10}{\second}}
\newcommand{\mnuc}{\SI{939}{\MeV}}
\newcommand{\mpi}{\SI{140}{\MeV}}

\begin{statement}{(Jackson 11.20)}
	The lambda particle ($\Lam$) is a neutral baryon of mass $M = \mLam$ that decays with a lifetime of $\tau = \tauLam$ into a nucleon of mass $\mq \approx \mnuc$ and a pi-meson of mass $\mw \approx \mpi$.  It was first observed in flight by its charged decay mode $\Lam \to p + \pi^-$ in cloud chambers.  The charged tracks originate from a single point and have the appearance of an inverted vee or lambda.  The particles' identities and momenta can be inferred from their ranges and curvature in the magnetic field of the chamber.
\end{statement}

\newcommand{\Pmu}{P^\mu}
\newcommand{\Psmu}{P_\mu}
\newcommand{\Pqmu}{{P_1}^\mu}
\newcommand{\Pqsmu}{{P_1}_\mu}
\newcommand{\Pwmu}{{P_2}^\mu}
\newcommand{\Pwsmu}{{P_2}_\mu}

\newcommand{\vp}{\vb{p}}
\newcommand{\vpq}{\vp_1}
\newcommand{\vpw}{\vp_2}

\begin{problem} \label{4.a}
	Using conservation of momentum and energy and the invariance of scalar products of 4-vectors show that, if the opening angle $\tht$ between the two tracks is measured, the mass of the decaying particle can be found from the formula
	\beq
		M^2 = \mq^2 + \mw^2 + 2 \Eq \Ew - 2 \pq \pw \cos\tht,
	\eeq
	where here $\pq$ and $\pw$ are the magnitudes of the 3-momenta.
\end{problem}

\begin{solution}
	The general momentum 4-vector for a particle is
	\beq
		\Pmu = (E / c, \vp),
	\eeq
	where $E$ is the energy of the particle and $\vp$ its three-dimensional momentum.  Let $\Pqmu$ and $\Pwmu$ be the momentum 4-vectors for the two particles, and define
	\beq
		\Pmu = \Pqmu + \Pwmu.
	\eeq
	Firstly, note that
	\beq
		\Pmu \Psmu = \frac{E^2}{c^2} - p^2.
	\eeq
	According to Jackson~(11.55),
	\beqn \label{E}
		E = \sqrt{c^2 p^2 + m^2 c^4},
	\eeqn
	so we have
	\beqn \label{thing4.1}
		\Pmu \Psmu = \frac{c^2 p^2 + M^2 c^4}{c^2} - p^2
		= M^2 c^2.
	\eeqn
	Note also that
	\beq
		\Pmu \Psmu = (\Pqmu + \Pwmu) (\Pqsmu + \Pwsmu)
		= \frac{(\Eq + \Ew)^2}{c^2} - (\vpq + \vpw)^2
		= \frac{\Eq^2 + 2 \Eq \Ew + \Ew^2}{c^2} - \pq^2 - 2 \vpq \vdot \vpw - \pw^2,
	\eeq
	and once again making use of \refeq{E},
	\beqn \label{thing4.2}
		\Pmu \Psmu = \frac{c^2 \pq^2 + \mq^2 c^4 + 2 \Eq \Ew + c^2 \pw^2 + \mw^2 c^4}{c^2} - \pq^2 - 2 \vpq \vdot \vpw - \pw^2
		= \mq^2 c^2 + \mw^2 c^2 + 2 \Eq \Ew / c^2 - 2 \pq \pw \cos\tht.
	\eeqn
	Equating \refeq{thing4.1} and \refeq{thing4.2}, we have
	\beq
		M^2 c^2 = \mq^2 c^2 + \mw^2 c^2 + 2 \Eq \Ew / c^2 - 2 \pq \pw \cos\tht.
	\eeq
	Taking $c = 1$, this becomes
	\beq
		M^2 = \mq^2 + \mw^2 + 2 \Eq \Ew - 2 \pq \pw \cos\tht
	\eeq
	as desired. \qed
\end{solution}



\newcommand{\ELam}{\SI{10}{\GeV}}

\begin{problem}
	A lambda particle is created with a total energy of {\ELam} in a collision in the top plate of a cloud chamber.  How far on the average will it travel in the chamber before decaying?  What range of opening angles will occur for a \SI{10}{\GeV} lambda if the decay is more or less isotropic in the lambda's rest frame?
\end{problem}

\begin{solution}
	According to Jackson~(11.51),
	\beq
		E = \gam m c^2.
	\eeq
	For the $\Lam$, this gives us $\gam = E / M$ in natural units.  Substituting this into the expression for time dilation, we find
	\beq
		\Del t = \gam \tau
		= \frac{E}{M} \tau
		= \frac{\ELam}{\mLam} (\tauLam)
		\approx \SI{2.60e-9}{\second}.
	\eeq

	\hl{opening angle?}
	
%	Let $\Pqmu$ represent the nucleon and $\Pwmu$ the pion.  Then from the result of \ref{4.a},
%	\beq
%		(\mLam)^2 = (\mnuc)^2 + (\mpi)^2 + 2 \Eq \Ew - 2 \pq \pw \cos\tht
%	\eeq
\end{solution}




\clearpage

\input{Problem5}



\vfill
In addition to the course lecture notes, I consulted Jackson's \emph{Classical Electrodynamics} while writing these solutions.
\end{document}