\documentclass[11pt]{article}
\usepackage{geometry, titlesec}
\usepackage[parfill]{parskip}
\usepackage[italicdiff]{physics}
\usepackage{amsfonts, amsthm}
\usepackage{fullpage}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{siunitx}
%\allowdisplaybreaks
 
%\renewcommand{\footrulewidth}{.2pt}
%\setlist[enumerate]{leftmargin=*}
\pagestyle{fancy}
\fancyhf{}
\lhead{Homework 2}
\rhead{Physics 133-B}
\setlength{\headheight}{11pt}
\setlength{\headsep}{11pt}
%\setlength{\footskip}{24pt}
%\cfoot{\today}
%\rfoot{\thepage}

\titleformat{\section}[runin]{\normalfont\large\bfseries}{Problem \thesection.}{1em}{}
\titleformat{\subsection}[runin]{\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subparagraph}[leftmargin]{\normalfont\normalsize\bfseries}{}{0pt}{}
\newcommand{\refeq}[1]{(\ref{#1})}

\newcommand{\beq}{\begin{equation*}}
\newcommand{\eeq}{\end{equation*}}

\newcommand{\beqn}{\begin{equation}}
\newcommand{\eeqn}{\end{equation}}

\newcommand{\qimplies}{\quad \implies \quad}


\newenvironment{statement}[1]
{
	\section{#1}
	\ignorespaces
}

\newenvironment{problem}
{
    \paragraph{Problem.}
    \color{darkgray}
    \ignorespaces
}

\newenvironment{solution}
{
    \paragraph{Solution.}
    \ignorespaces
}

\begin{document}

\newcommand{\dist}{\SI{2}{\meter}}
\newcommand{\freq}{\SI{800}{\Hz}}
\newcommand{\vsound}{\SI{344}{\meter\per\second}}
	
\begin{itemize}
% Y&F 16.61
\item Consider two speakers emitting sound at the same volume with frequency $f = \freq$.  One speaker is located at the origin, and the other on the $y$ axis at $y = \dist$.  At what locations on the positive $x$ axis is the interference completely constructive?  At what points is it completely destructive?
	
Now we decrease $f$ until there are no longer any points of completely destructive interference on the positive $x$ axis.  How low must $f$ be for this to occur?


\bigskip\bigskip\bigskip

\newcommand{\ofreq}{\SI{10.5}{\giga\Hz}}
\newcommand{\ifreq}{\SI{12.5}{\giga\Hz}}
\newcommand{\pspeed}{\SI{35}{\meter\per\second}}


% Y&F 16.51
\item A police car is waiting on the shoulder of Lake Shore Drive.  In order to catch speeding commuters, the officer uses a radar gun that emits sound of frequency {\ofreq}.  There is a single speeding car on the road, which is directly in front of or behind her.  When she fires the gun, the frequency that returns from the speeder's car is {\ifreq}.  What is the speed of the car?  Is it moving toward the officer, or away from her?

A few moments later, the officers takes off at {\pspeed} in pursuit of the speeder, who does not change his speed.  If the officer fires the gun now, what frequency will she receive?

\end{itemize}




\end{document}