\documentclass[11pt]{article}
\usepackage{homework}

\classname{443}
\homeworknum{1}


\begin{document}

\input{definitions}
\input{probdefs}


\state{(Peskin \& Schroeder 2.1)}{
	Classical electromagnetism (with no sources) follows from the action
	\al{
		S &= \int \ddqx \paren{ -\frac{1}{4} \Fsmn \Fmn }, &
		\where \Fsmn = \ptsm \Asn - \ptsn \Asm.
	}
}



\prob{}{
	Derive Maxwell's equations as the Euler-Lagrange equations of this action, treating the components $\Asm(x)$ as the dynamical variables.  Write the equations in standard form by identifying $\Ei = -\Foi$ and ${\epsijk \Bk = -\Fij}$.
}



\prob{}{
	Construct the energy-momentum tensor for this theory.  Note that the usual procedure does not result in a symmetric tensor.  To remedy that, we can add to $\Tmn$ a term of the form $\ptsl \Klmn$, where $\Klmn$ is antisymmetric in its first two indices.  Such an object is automatically divergenceless, so
	\eq{
		\Thmn = \Tmn + \ptsl \Klmn
	}
	is an equally good energy-momentum tensor with the same globally conserved energy and momentum.  Show that this construction, with
	\eq{
		\Klmn = \Fmn \An,
	}
	leads to an energy-momentum tensor $\Th$ that is symmetric and yields the standard formulae for the electromagnetic energy and momentum densities:
	\al{
		\cE &= \frac{E^2 + B^2}{2}; &
		\vS &= \vE \cross \vB.
	}
}



\state{The complex scalar field (Peskin \& Schroeder 2.2)}{
	Consider the field theory of a complex-valued scalar field obeying the Klein-Gordon equation.  The action of this theory is
	\eq{
		S = \int \ddqx (\ptsm \phis \ptm \phi - m^2 \phis \phi).
	}
	It is easiest to analyze this theory by considering $\phix$ and $\phisx$, rather than the real and imaginary parts of $\phix$, as the basic dynamical variables.
}



\prob{}{
	Find the conjugate momenta to $\phix$ and $\phisx$ and the canonical commutation relations.  Show that the Hamiltonian is
	\eq{
		H = \int \ddcx (\pis \pi + \grad \phis \vdot  \grad \phi + m^2 \phis \phi).
	}
	Compute the Heisenberg equation of motion for $\phix$ and show that it is indeed the Klein-Gordon equation.
}



\prob{}{
	Diagonalize $H$ by introducing creation and annihilation operators.  Show that the theory contains two sets of particles of mass $m$.
}



\prob{}{
	\label{2.3(c)}
	
	Rewrite the conserved charge
	\eq{
		Q = \int \ddcx \frac{i}{2} (\phis \pis - \pi \phi)
	}
	in terms of creation and annihilation operators, and evaluate the charge of the particles of each type.
}



\prob{}{
	Consider the case of two complex Klein-Gordon fields with the same mass.  Label the fields as $\phiax$, where $a = 1, 2$.  Show that there are now four conserved charges, one given by the generalization of part~\ref{2.3(c)}, and the other three given by
	\eq{
		\Qi = \int \ddcx \frac{i}{2} (\phias \sigiab \pibs - \pia \sigiab \phib),
	}
	where $\sigi$ are the Pauli sigma matrices.  Show that these three charges have the commutation relations of angular momentum ($SU(2)$).  Generalize these results to the case of $n$ identical complex scalar fields.
}




\state{(Peskin \& Schroeder 2.3)}{
	\label{3}
	
	Evaluate the function
	\eq{
		\ev{\phix \, \phiy}{0} = D(x - y)
	= \int \ddpf \frac{1}{2 \Ep} e^{i p \, (x - y)},
	}
	for $(x - y)$ spacelike so that $(x - y)^2 = -r^2$, explicitly in terms of Bessel functions.
}





\state{}{
	The classical limit of a harmonic oscillator can be described in terms of \emph{coherent states},
	\eq{
		\kalp = \exp( \alp \adag - \frac{1}{2} |\alp|^2 ) \ko.
	}
	When $\alp$ is large, the oscillator state is semiclassical.  Proceeding similarly for the Fourier modes of the quantum Klein-Gordon field,
	\al{
		\kf &= \Nf \exp( i \int \ddpf \fp \, \apdag ) \ko, &
		\Nf &= \exp( -\frac{1}{2} \int \ddpf | \fp |^2 ).
	}
}



\prob{}{
	Evaluate the expectation value of the field operator $\ev{\phix}{f}$ and show that it satisfies the Klein-Gordon equation.
}



\prob{}{
	Evaluate the relative mean square fluctuation of the occupation number of the mode with momentum $\vp$ and the relative mean square fluctuation in the total energy:
	\al{
		&\frac{\evnhps - \evnhp^2}{\evnhp^2}, &
		&\frac{\evHs - \evH^2}{\evH^2}.
	}
	Is either of these a good measure of the degree to which the field is classical?  Justify your answer.
}



\prob{}{
	Take $\Delta(x - y) = \ev{\phivx \, \phivy}{0}$ (equal times) as a measure of the fluctuations or correlations of the field amplitude.  Use your result from problem~\ref{3} to evaluate this quantity.  What is the meaning of the divergence as $\vx \to \vy$?
}




%\input{Problem5}


%\makebib

\end{document}
